Hijacking Simulators with Universal Probabilistic Programming

### 2019-08-14 paper/learning/notes.tex
stuff to ask tom

In addition to this, if we have direct access to a rejection sampling loop then it is indeed 
possible to construct the density, up to a constant of proportionality. 
That is:
\begin{align*}
y \sim q(\theta) \\
do something \\
\alpha = p_{accept}(y,\theta) \\
\mathtt{if} Bern(\alpha) = 1: \\
return y
\end{align*}
thus the density of $y,\theta \propto q(\theta, y) \alpha$.
However, in order to do construct this we need the input $\theta$, $y$, $\alpha$ and the distributions utilised for the rejection sampling loop. 
For any arbitrary simulator, written in a generic code base, this is an incredibly hard 
task to do, because rejection sampling blocks within simulators are not necessarily written in a standard template
\textbf{Add examples in a supplement.  Need to demonstrate to YW that this is challenging to do, even if you know where 
the rejection sampling loops are. Hypothesis, RS loops are not written in a templated form}. With the PPX protocols previously developed in \cite{baydin2018efficient,gram2019hijacking} we can however, 
extract all variables, other than $\alpha$ in an automated way.
However, without being able to easily extract $\alpha$ this means that we are unable to 
correctly build the form of the density and thus we limited in our ability to leverage 
powerful inference methods such as MCMC and Variational inference.
Thus, we propose another approach, automating the amortization of rejection sampling loops. 
To learn functions that we can easily sample from and use to construct the density of the entire
simulator.

@@

Thus, to define the objective for each subproblem $r  \in S_{R}$ as:
\begin{equation}
  J_{r}(\eta_r) = -\mathbb{E}_{p(x)}[\sum^{n_x}_{j=1}\mathbb{I}(a_j = r)\ln(q_r(x_j;\eta_r, \phi_j))]
\end{equation}

@@

\paragraph{TODO list}
\begin{itemize}
  \item Implement baselines
  \item Implement plotting scripts and evaluation scripts
  \item testing code and infrastructure
  \item litrature review on any methods to learn rejection sampling methods, 
  and conditional density estimation. See this paper for kernel approximation 
  methods~\cite{erraqabi2016pliable}. In addition to the methods listed in the 
  rejection sampling subsection. 
\end{itemize}
\subsection{Ground truth}

@@

\paragraph{observations} As the functions are relatively simple the loss functions it gets t
to a loss of around 0.6 and stops decreasing. Although, this seems low, I don't think that it is
actually good, because that would mean on average, for each data point there is a loss of $\pm$ 0.6
which per data point is quite bad. That may be for a number of reasons, using MSE loss, using a simple
model architecture. It really struggles to model the tails. It may require a much larger batch input size. 
. Increasing the batch size helped to correct for the peaked density, however the behavior is still odd.
The peaked behaviour still does not disappear, this, I believe is due to the loss function averaging 
over all points,

@@

\begin{itemize}
  \item Isn't what we are doing just conditional density estimation? It seems
  like there is a lot of prior work here. I.e see this paper here for a good overview ~\cite{greenberg2019automatic}
  \item It does seem like this field has a large extensive past, but we still have a small neesh .
\end{itemize}

>>>>

In addition to this, if we have direct access to a rejection sampling loop then it is indeed 
possible to construct the density, up to a constant of proportionality. 
That is:
\begin{align*}
y \sim q(\theta) \\
do something \\
\alpha = p_{accept}(y,\theta) \\
\mathtt{if} Bern(\alpha) = 1: \\
return y
\end{align*}
thus the density of $y,\theta \propto q(\theta) \alpha$.
However, in order to do construct this we need the input $\theta$, $y$, $\alpha$ and the distributions utilised for the rejection sampling loop. 
For any arbitrary simulator, written in a generic code base, this is an incredibly hard 
task to do, because rejection sampling blocks within simulators are not necessarily written in a standard template
\textbf{Add examples in a supplement.  Need to demonstrate to YW that this is challenging to do, even if you know where 
the rejection sampling loops are. Hypothesis, RS loops are not written in a templated form}. With the PPX protocols previously developed in \cite{baydin2018efficient,gram2019hijacking} we can however, 
extract all variables, other than $\alpha$ in an automated way.
However, without being able to easily extract $\alpha$ this means that we are unable to 
correctly build the form of the density and thus we limited in our ability to leverage 
powerful inference methods such as MCMC and Variational inference.
Thus, we propose another approach, automating the amortization of rejection sampling loops. 
To learn functions that we can easily sample from and use to construct the density of the entire
simulator.

@@

Thus, to define the objective for each subproblem $r  \in S_{R}$ as:
\begin{equation}
  \label{eq:objective}
  J_{r}(\eta_r) = -\mathbb{E}_{p(x)}[\sum^{n_x}_{j=1}\mathbb{I}(a_j = r)\ln(q_r(x_j;\eta_r, \phi_j))]
\end{equation}

@@

\paragraph{TODO list}
\begin{itemize}
  \item Implement baselines
  \item Implement plotting scripts and evaluation scripts
  \item testing code and infrastructure
  \item litrature review on any methods to learn rejection sampling methods, 
  and conditional density estimation. See this paper for kernel approximation 
  methods~\cite{erraqabi2016pliable}. In addition to the methods listed in the 
  rejection sampling subsection. 
\end{itemize}

\paragraph{Tom Questions}

\begin{itemize}
  \item I don't understand how the objective function is comparing the generated samples from the density 
  estimator to the ground truth, at least in the way it is currently defined? 
  \item All RS references I look at state that we must sample from a uniform distribution and
  then check if our proposed value is less than that uniformly sampled point, yet our RS do not have this,
  nor what is the acceptance probability within our rejection sampling statements? The proof for rejection sampling seems to rely on the fact that you are drawing from a uniform distribution.

  \end{itemize}
\subsection{Ground truth}

@@

\paragraph{observations} As the functions are relatively simple the loss functions it gets t
to a loss of around 0.6 and stops decreasing. Although, this seems low, I don't think that it is
actually good, because that would mean on average, for each data point there is a loss of $\pm$ 0.6
which per data point is quite bad. That may be for a number of reasons, using MSE loss, using a simple
model architecture. It really struggles to model the tails. It may require a much larger batch input size. 
. Increasing the batch size helped to correct for the peaked density, however the behavior is still odd.
The peaked behaviour still does not disappear, this, I believe is due to the loss function averaging 
over all points.

@@

\begin{itemize}
  \item How is Equation~\ref{eq:}
  \item Isn't what we are doing just conditional density estimation? It seems
  like there is a lot of prior work here. I.e see this paper here for a good overview ~\cite{greenberg2019automatic}
  \item It does seem like this field has a large extensive past, but we still have a small niche.
\end{itemize}
